---
title: "viz_part2"
author: "Eunsuh Cho"
date: "2023-10-03"
output: github_document
---

Loading the necessary packages

```{r}
library(tidyverse)
library(patchwork)
```

Loading the NOAA weather data using last week's code

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

weather_df
```

Scatterplot - tmax against tmin

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .5)
```

Labels

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temprature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package"
  )
```

Scales

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temprature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package") +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15C", "0", "15"))
```

Other `scale_x_*` and `scale_y_*` options

``` {r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temprature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package") +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15C", "0", "15"),
    limits = c(-20, 30)) +
  scale_y_continuous(
    trans = "sqrt",
    position = "right")
```
`scale_y_sqrt()` can transform the Y scale, and `xlim()` can control X axis plot limits.

Controlling `color` aesthetic using `scale`s

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temprature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package") +
  scale_color_hue(name = "Location", h = c(100, 300))
```

Using `viridis` package

```{r}
ggp_temp_plot =
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temprature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package") +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  )

ggp_temp_plot
```
We used `discrete = TRUE` bceause the `color` aesthetic is mapped to a discrete variable!

Themes

`theme_gray`
```{r}
ggp_temp_plot +
  theme(legend.position = "bottom")
```

`theme_bw()`
```{r}
ggp_temp_plot +
  theme_bw() +
  theme(legend.position = "bottom")
```

`theme_classic()`
```{r}
ggp_temp_plot +
  theme_classic() +
  theme(legend.position = "bottom")
```

lol excel 2003 theme
```{r}
ggp_temp_plot +
  ggthemes::theme_excel() +
  theme(legend.position = "bottom")
```

Learning assessment
```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth(se = FALSE) +
  geom_point(aes(size = prcp), alpha = .75) +
  labs(
    title = "Temperature plot",
    x = "Date",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package") +
  viridis::scale_color_viridis(discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "bottom")
```





