---
title: "viz_part2"
author: "Eunsuh Cho"
date: "2023-10-03"
output: github_document
---

Loading the necessary packages

```{r}
library(tidyverse)
library(patchwork)
```

Loading the NOAA weather data using last week's code

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

weather_df
```

Scatterplot - tmax against tmin

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .5)
```

## Labels

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temprature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package"
  )
```

## Scales

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temprature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package") +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15C", "0", "15"))
```

Other `scale_x_*` and `scale_y_*` options

``` {r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temprature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package") +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15C", "0", "15"),
    limits = c(-20, 30)) +
  scale_y_continuous(
    trans = "sqrt",
    position = "right")
```
`scale_y_sqrt()` can transform the Y scale, and `xlim()` can control X axis plot limits.

Controlling `color` aesthetic using `scale`s

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temprature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package") +
  scale_color_hue(name = "Location", h = c(100, 300))
```

Using `viridis` package

```{r}
ggp_temp_plot =
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temprature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package") +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  )

ggp_temp_plot
```
We used `discrete = TRUE` bceause the `color` aesthetic is mapped to a discrete variable!

### Themes

`theme_gray`
```{r}
ggp_temp_plot +
  theme(legend.position = "bottom")
```

`theme_bw()`
```{r}
ggp_temp_plot +
  theme_bw() +
  theme(legend.position = "bottom")
```

`theme_classic()`
```{r}
ggp_temp_plot +
  theme_classic() +
  theme(legend.position = "bottom")
```

lol excel 2003 theme
```{r}
ggp_temp_plot +
  ggthemes::theme_excel() +
  theme(legend.position = "bottom")
```

Learning assessment
```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth(se = FALSE) +
  geom_point(aes(size = prcp), alpha = .75) +
  labs(
    title = "Temperature plot",
    x = "Date",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package") +
  viridis::scale_color_viridis(discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

figure preferences in global options; copy and paste this at the beginning of every new file

```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Data argument in `geom_*`

```{r}
central_park_df =
  weather_df |> 
  filter(name == "CentralPark_NY")

molokai_df =
  weather_df |> 
  filter(name == "Molokai_HI")

ggplot(data = molokai_df, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park_df)
```

## patchwork

```{r}
tmax_tmin_p =
  weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p =
  weather_df |> 
  filter(prcp > 0) |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .5) +
  theme(legend.position = "none")

tmax_date_p =
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```

## Data manipulation

Reordering `name` by hand

```{r}
weather_df |> 
  mutate(name = forcats::fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_violin(aes(fill = name), color = "blue", alpha = .5) +
  theme(legend.position = "bottom")
```

Reordering `name` according to `tmax` values in each name

```{r}
weather_df |> 
  mutate(name = forcats::fct_reorder(name, tmax)) |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_violin(aes(fill = name), color = "blue", alpha = .5) +
  theme(legend.position = "bottom")
```

Making a 3 panel plot showing densities for `tmax` and `tmin` within each location

```{r}
weather_df |> 
  select(name, tmax, tmin) |> 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temp") |> 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = .5) +
  facet_grid(~name) +
  viridis::scale_fill_viridis(discrete = TRUE)
```

Emphasizing data tidiness for visualization

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") |>
  select(id, visit, everything()) |>
  mutate(
    visit = recode(visit, "bl" = "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) |>
  arrange(id, visit)

ggplot(pulse_data, aes(x = visit, y = bdi)) + 
  geom_boxplot()
```

Revisiting FAS data

```{r}
pup_data = 
  read_csv("data/FAS_pups.csv") |>
  janitor::clean_names() |>
  mutate(
    sex = 
      case_match(
        sex, 
        1 ~ "male", 
        2 ~ "female"))

litter_data =
  read_csv("data/FAS_litters.csv") |> 
  janitor::clean_names() |> 
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pup_data, litter_data, by = "litter_number")

fas_data |> 
  select(sex, dose, day_of_tx, pd_ears:pd_walk) |> 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome",
    values_to = "pn_day") |> 
  drop_na() |> 
  mutate(outcome = forcats::fct_reorder(outcome, pn_day, median)) |> 
  ggplot(aes(x = dose, y = pn_day)) +
  geom_violin() +
  facet_grid(day_of_tx ~ outcome)
```






